---
title: "Extracting environmental data with `rvdat`"
date: 2024-01-10
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting environmental data with `rvdat`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

We often want to extract the environmental data that some of our receivers log. You can do this by going into Innovasea's VUE or Fathom Connect programs and clicking a bunch of buttons, or you can do it programmatically and reproducibly by combining `matos` and [`rvdat`](https://rvdat.obrien.page).

```{r setup}
library(matos)
library(rvdat)
```

## Download the data

First, log into your MATOS account.

```{r}
matos_login()
```

Next, tell `rvdat` the location of the `vdat.exe` executable (check out the [`rvdat` documentation](https://rvdat.obrien.page) for more information on this).

```{r, eval=FALSE}
vdat_here('vdat.exe')
```

Using `matos`, I'm going to find the VDAT files associated with the [Mid-Bay Chesapeake Backbone array](https://matos.asascience.com/project/detail/161). Note that things like temperature and tilt are logged within the VRL/VDAT files, and so will be under the "detections" file type.

```{r}
detection_files <- list_project_files(
  project = 'UMCES Chesapeake Backbone, Mid-Bay',
  file_type = 'detections'
)

head(detection_files)
```

I'll download the first file into a temporary directory.

```{r}

get_project_file(
  url = detection_files$url[1],
  out_dir = tempdir()
)

vrl_file <- list.files(tempdir(), pattern = 'vrl$', full.names = T)
```

## Convert VRL using `rvdat`

Now, I'll use `rvdat::vdat_to_folder` to convert that file into a folder of CSV files, each one representing one data type.

```{r}
vdat_to_folder(
  vrl_file,
  outdir = tempdir()
) 
```

There are quite a few files in there, but for this we're going to focus on the temperature records stored in `TEMP.csv`.

```{r}
list.files(tempdir(), pattern = 'fathom-split', full.names = T) |> 
  list.files()

bwt_file <- list.files(tempdir(), pattern = 'TEMP', full.names = T, recursive = T)
```

## Read it into R

Let's read in the data.

```{r, error=TRUE}
read.csv(bwt_file)
```

An error, oh no! Let's see what's causing it.

```{r}
read.csv(bwt_file,
         header = FALSE,
         nrows = 5)
```

Ah, the data doesn't really start until the second row. Skip the first one and take a look.

```{r}
bwt <- read.csv(
  bwt_file,
  skip = 1
)

head(bwt)
```

Great! The data are in. I'm going to clean up the names a little bit and convert the time column from a character string to POSIX time.

```{r}
names(bwt) <- gsub('[_\\.]', '', tolower(names(bwt)))

names(bwt)

bwt$devicetimeutc <- as.POSIXct(bwt$devicetimeutc,
                                tz = 'UTC',
                                format = '%F %T')

head(bwt)
```

## See what we've got

Now that we have our time series, let's see what it looks like!

```{r}
plot(ambientdegc ~ devicetimeutc,
     data = bwt,
     type = 'l')

```

## Other variables

We can pull out other variables in a similar manner. Take, for example, receiver tilt, located in `ATTITUDE.csv`.

```{r}
env_import <- function(file){
  hold <- list.files(
    tempdir(),
    pattern = file,
    full.names = T,
    recursive = T
  ) |> 
    read.csv(skip = 1)
  
  names(hold) <- gsub('[_\\.]', '', tolower(names(hold)))
  
  hold$devicetimeutc <- as.POSIXct(hold$devicetimeutc,
                                   tz = 'UTC',
                                   format = '%F %T')
  
  hold
}

tilt <- env_import('ATTITUDE')

plot(tiltdeg ~ devicetimeutc, data = tilt, type = 'l')
```

Depth, in `DEPTH.csv`:

```{r}
depth <- env_import('DEPTH')

plot(depthm ~ devicetimeutc, data = depth)
```

Battery, in `BATTERY.csv`:

```{r}
battery <- env_import('BATTERY')

plot(batteryvoltagev ~ devicetimeutc, data = battery)
```

Noise, which is actually in the diagnostic file `DIAG.csv`:

```{r}
diagnostics <- env_import('DIAG')

plot(noisemeanmv ~ devicetimeutc, data = diagnostics)
```
Hourly summaries of pings and detections are also in the diagnositc file:

```{r}
plot(ppmdetections ~ devicetimeutc, data = diagnostics)
plot(ppmpings ~ devicetimeutc, data = diagnostics)
```

But are really located in `DET.csv`.

```{r}
dets <- env_import('DET')

head(dets[, ])
```

These data were gleaned from a particular receiver programmed in a particular way -- note that many of the data fields are empty! Depending on the receiver, how you programmed it, and the arguments you passed to `rvdat` (time correction, anyone?), your split CSV could look rather different.
