---
title: "Summarizing with otndo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summarizing with otndo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(matos)
```

## ACT members

If you're an ACT member, the first thing you'll want to do is list the projects for which you have access.

```{r}
list_my_projects(read_access = T)
```

From here, you can just provide your project name or number and let it rip. Below, I provide the number associated with "UMCES BOEM Offshore Wind Energy".

```{r, eval=FALSE}
matos_receiver_summary(87)
```


This function does a bit of data cleaning on the front end and then runs everything through a [Quarto](https://quarto.org/) or [RMarkdown](https://rmarkdown.rstudio.com/) report. The functions use Quarto by default, but RMarkdown will be selected if:

1.  Quarto is not installed on the computer, or
2.  the `rmd` argument is set to `TRUE`.

```{r, eval=FALSE}
# Compiles with Quarto (default)
matos_receiver_summary(87, rmd = F)

# Compiles with RMarkdown
matos_receiver_summary(87, rmd = T)
```

Functionality is identical for `matos_tag_summary`:

```{r, eval=FALSE}
matos_tag_summary(87)
```
